---
title: 'Project: CO2, Carbon Dioxide Uptake in Grass Plants'
author: "Carayon Chloé - Taillieu Victor"
date: "05/11/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Description of the dataset

The CO2 data frame is the result of an experiment based on the cold tolerance of the grass species Echinochloa crus-galli.
We load the data and get some information about it thanks to this commands:

```{r description}
data(CO2)
dim(CO2)
names(CO2)
```

The CO2 dataset is composed of 84 rows ans 5 columns of names
"Plant", "Type", "Treatment", "conc" and "uptake".
Using the help section, we learn that:

* Plant is a factor giving to each plant a unique identifier:
  * Q or M represents the origin of the plant, respectively Quebec or Mississippi
  * n or c gives information about the treatment
  * 1, 2 or 3 is the id of the plant

For example, Mc1 is plant 1 from Mississippi and was chilled.

* Type is a factor with levels Quebec and Mississippi
* Treatment is a factor with levels chilled and nonchilled
* conc is the ambient carbon dioxide concentrations' vector, measured in mL/L
* uptake is the carbon dioxide uptake rates' vector in umol/m² sec

The CO2 uptake rate of six plants from Quebec and six plants from Mississippi was measured at several levels of ambient CO2 concentration. 
Half the plants were chilled overnight before the experiment was conducted.

## Data exploration

### Summary

Let's display the dataset and its structure:

```{r structure}
CO2
class(CO2)
str(CO2)
```

The output confirms what we have seen before thanks to the functions and help section.

This dataset was originally part of another package (nlme) which has its own grouped-data classes.
That is why there are some unusual classes like nfnGroupedData, nfGroupedData, groupedData and attr.

```{r summary}
summary(CO2)
levels(CO2$Plant)
```

Thanks to the summary, we obtain information about each column. We execute the levels function on CO2$Plant as we don't have all the information displayed in the summary. 
We see that 3 plants were put in several conditions (with their name corresponding to the type, treatment and number) and measurements were made on their carbon dioxide uptake rates according to the concentration of ambient carbon dioxide.

By grouping plants by their type (Quebec - Mississipi), we can see some characteristics:

```{r tapply1}
tapply(CO2$uptake, CO2$Type, summary)
```

Plants from Quebec have an uptake rate between 9.30 and 45.50 whereas for Mississippi, it is between 7.70 and 35.50.
Moreover, for Quebec, half of plants have an uptake rate greater than 37.15 while for Mississippi, it is 19.30.
There are huge differences between Quebec and Mississippi minimums, maximums, medians and also quartiles.
A quarter of Quebec plants have an uptake rate under 30.32 while for Mississipi it is 13.88.
Finally, for 75% of Mississippi plants, the uptake rate is less than 28.05 whereas for Quebec it is less than 40.15.

So it appears that Quebec plants seem to have higher uptake rates than Mississippi plants.

Let's take a look on two plants from Quebec, one chilled and one non chilled:

```{r tapply2}
tapply(CO2$uptake, CO2$Plant, summary)
```

Let's take the examples of Qn3 and Qc3:

- Qn3 (third plant species of Quebec nonchilled)
have an uptake rate between 16.20 and 45.50 with an average of 37.61.
Half of the Qn3 have a rate less than 42.10.
In those plants, 25% have an uptake rate less than 36.35.
Moreover, 75% have a rate less than 43.40.
In fact 50% of the population uptake rate is between 36.35 and 43.40 (mL/L).

- Qc3 (third plant species of Quebec chilled):
have an uptake rate between 15.10 and 41.40 with an average of 32.59.
Half of the Qc3 have a rate less than 38.10.
In those plants, 25% have an uptake rate less than 27.50.
Moreover, 75% have a rate less than 39.25.
In fact 50% of the population uptake rate is between 27.50 and 39.25 (mL/L).

So we see that for two plants numbered 3 from the same origin (Quebec) but one chilled and one non chilled, we obtain a significant difference in uptake rate.

We can sum up these differences using the summarize function from dplyr library.

```{r summarize}
library(dplyr)
summarize(group_by(CO2, Type), average_uptake = mean(uptake))
summarize(group_by(CO2, Treatment), average_uptake = mean(uptake))
summarize(group_by(CO2, conc), average_uptake = mean(uptake))
```

From these results, we observe that the average uptake for plants from Quebec is much higher than the one for plants from Mississippi (33.5 and 20.9 umol/m² sec respectively).
If we compare them according to the treatment, chilled plants have a smaller average uptake compared to nonchilled ones (23.8 as opposed to 30.6 umol/m² sec).
Finally, it makes sense to see that when we increase the concentration of ambient carbon dioxide, the plants tend to have higher carbon dioxide uptake rates.

## Data visualization

### Global overview

To begin with, let's just plot the data without any filtering:

```{r}
library(ggplot2)
qplot(conc, uptake, data = CO2, geom = "point",
      main = "CO2 uptake rate according to ambient CO2 concentration",
      xlab = "CO2 concentration (mL/L)", ylab = "CO2 uptake rate (umol/m² sec)")
```

The only thing we can say is that CO2 uptake rates seem to increase as ambient CO2 concentration increase.

Then, we can separate plants by their characteristics to get some insights:

```{r}
qplot(conc, uptake, facets = Treatment ~ ., data = CO2, color = Type,
      geom = c("point", "smooth"), se = FALSE,
      main = "CO2 uptake according to ambient CO2 concentration",
      xlab = "CO2 concentration (mL/L)", ylab = "CO2 uptake rate (umol/m² sec)")
```

Here we can observe several interesting things:

- As we noticed in the previous part, plants from Quebec (in red) almost always have a higher CO2 uptake rate compared to plants from Mississippi for the same ambient concentration. The smooth tendency lines clearly show that.
- Moreover, nonchilled plants (in the upper graph) seem to jump higher in terms of uptake rates that chilled ones.

We can say that this visualization confirms what we pointed out in the data exploration part.
Now, let's go further by analyzing the distribution of uptake rates according to the country of origin or treatment.

```{r}
qplot(uptake, data = CO2, color = Type, geom = "density",
      main = "Plants CO2 uptake distribution according to their origin",
      xlab = "CO2 uptake rate (umol/m² sec)")
qplot(uptake, data = CO2, color = Treatment, geom = "density",
      main = "Plants CO2 uptake distribution according to their treatment",
      xlab = "CO2 uptake rate (umol/m² sec)")
```

Again, these two plots confirm our hypothesis made above.
An important part of Mississippi plants have an uptake rate between 15 and 20 umol/m² sec whereas for Quebec plants the peak is at about 40 umol/m² sec.
For the treatment, a big part of chilled plants have an uptake rate between 15 and 20 umol/m² sec whereas for nonchilled plants the peak is higher, at about 30 umol/m² sec.

### Boxplot of uptake by origin

Now, we are going to use boxplots to visualize the spreading of the uptake rates.
We can group our plants by types (Quebec and Mississippi):

```{r}
ggplot(CO2, aes(x=Type, y=uptake,  fill=Type)) +
  geom_boxplot(show.legend = FALSE) +
  ggtitle("Plants CO2 uptake rates depending on their origin") +
  xlab("Origin") + ylab("CO2 uptake rate (umol/m² sec)")
```

This figure confirms what we saw in the data exploration part. There is a huge difference in the uptake rates between plants from Quebec and plants from Mississippi. As we know 50% of the data is in the box and by visualizing those boxplots, we clearly see that on average, plants from Quebec have higher uptake rates than Mississippi plants. In fact 25% of Quebec plants have less than 30 for their uptake rate whereas more than 75% of Mississippi plants are under 30 for their uptake rates.

### Boxplots of uptake depending on treatment

Now, let's take a look at the importance of the treatment on those plants.
Let's compare the uptake rate depending on the treatment.

```{r}
ggplot(CO2, aes(x=Treatment, y=uptake,  fill=Treatment)) +
  geom_boxplot(show.legend = FALSE) +
  ggtitle("Plants CO2 uptake rates depending on treatment") +
  xlab("Treatment") + ylab("CO2 uptake rate (umol/m² sec)")
```

Now, if we combine our plots, we will obtain the following result:

```{r}
ggplot(CO2, aes(x=Type, y=uptake, fill=Treatment)) + geom_boxplot() +
  ggtitle("CO2 uptake rates of plants depending on origin with treatment specified") +
  xlab("Origin") + ylab("CO2 uptake rate (umol/m² sec)")
```

* For Quebec

In the previous part, we studied two plants from Quebec with different treatments and find out that there was a significant difference in their uptake rates. It appears clearly here as less than 75% of chilled plants are under 39 while half of nonchilled have an uptake rate of more than 39.

* For Mississippi

Those differences are even more relevant with Mississippi plants:
nonchilled Mississippi plants have box where more than 25% are upper 22 whereas for chilled plants, less than 75% of them can reach 19.

Thanks to those boxplots, we can clearly see that depending on the type of plants and the treatment, there are some chance to obtain very different uptake rates.

## Testing hypothesis

Let's work on hypothesis regarding the treatment and the type.

### Hypothesis on treatment 

Firstly, we make an assumption on the treatment of the plants. 
Our null hypothesis is that treatment on plant will not affect the uptake rates.
Our alternative hypothesis is that treatment on plants may affect the uptake rates.

#### P value, one way test 

```{r}
t.test(CO2$uptake~CO2$Treatment)
```

Looking on p-value for treatment, which is 3.107ee-3, we see that it is smaller than the alpha value 0.05.
So, we can reject our null hypothesis and conclude that the treatment may make a difference on the uptake rates.

### Hypothesis on Type 

Secondly, we make an assumption on the type of the plants. 
Our null hypothesis is that type will not affect the uptake rates
Our alternative hypothesis is that type of plants may affect the uptake rates.

#### P value, one way test 

```{r}
t.test(CO2$uptake~CO2$Type)
```

Looking on p-value for type, which is 4.451e-9, we see that it is smaller than the alpha value 0.05.
So, we can reject our null hypothesis and conclude that the type may make a difference on the uptake rates.

### Combined our hypothesis

We combine our two hypothesis and make an assumption that the relationship between Type and Treatment will not make a difference on the uptake rates.
By combining our hypothesis:

#### Correlation 
```{r}
cor.test(CO2$uptake,CO2$conc)
```


#### Linear Model 

```{r}
Treatm_Type_fit = lm(CO2$uptake ~ CO2$Type*CO2$Treatment)
anova(Treatm_Type_fit)

```

We obtain 2.378e-10 for Type and 1.817e-4 for Treatment for p-value. Those values are smaller than 0.05, so we can again reject the null hypothesis and conclude that Type and Treatment may make a difference on the uptake rates.
But looking at the p-value for relationship between Treatment and Type, we obtain 0.0642 which is bigger than 0.05. So, we can't reject the null hypothesis and we conclude that the relationship between Type and Treatment did not seriously impact the uptake rates.


## Fitting model
