---
title: 'Project: CO2, Carbon Dioxide Uptake in Grass Plants'
author: "Carayon Chloé - Taillieu Victor"
date: "05/11/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Description of the dataset

The CO2 data frame is the result of an experiment based on the cold tolerance of the grass species Echinochloa crus-galli.
We load the data and obtain some information thanks to those commands:

```{r description}
data(CO2)
dim(CO2)
names(CO2)

CO2[sample(nrow(CO2)),]
```

The CO2 dataset is composed of 84 rows ans 5 columns of names
"Plant", "Type", "Treatment", "conc" and "uptake".
Using the help section, we learn that:

* Plant is a factor giving to each plant a unique identifier:
  * Q or M represents the origin of the plant, respectively Quebec or Mississippi
  * n or c gives information about the treatment
  * 1, 2 or 3 is the id of the plant

For example, Mc1 is plant 1 from Mississippi and was chilled.

* Type is a factor with levels Quebec and Mississippi
* Treatment is a factor with levels chilled and nonchilled
* conc is the ambient carbon dioxide concentrations' vector, measured in mL/L
* uptake is the carbon dioxide uptake rates' vector in umol/m² sec

The CO2 uptake rate of six plants from Quebec and six plants from Mississippi was measured at several levels of ambient CO2 concentration. 
Half the plants were chilled overnight before the experiment was conducted.

## Data exploration

### Summary

Let's display the dataset and its structure:

```{r structure}
CO2
class(CO2)
str(CO2)
```

The output confirms what we have seen before thanks to the functions and help section.

This dataset was originally part of another package (nlme) which has its own grouped-data classes.
That is why there are some unusual classes like nfnGroupedData, nfGroupedData, groupedData and attr.

```{r summary}
summary(CO2)
levels(CO2$Plant)
```

Thanks to the summary, we obtain information about each column. We execute the levels function on CO2$Plant as we don't have all the information displayed in the summary. 
We see that 3 plants were put in several conditions (with their name corresponding to the type, treatment and number) and measurements were made on their carbon dioxide uptake rates according to the concentration of ambient carbon dioxide.

### Characteristics

By grouping plants by their type (Quebec - Mississipi), we can see some characteristics:

```{r tapply1}
tapply(CO2$uptake, CO2$Type, summary)
```

Plants from Quebec have an uptake rate between 9.30 and 45.50  umol/m² sec whereas for Mississippi, it is between 7.70 and 35.50  umol/m² sec.
Moreover, for Quebec, half of plants have an uptake rate greater than 37.15 while for Mississippi, it is 19.30.
There are huge differences between Quebec and Mississippi minimums, maximums, medians and also quartiles.
A quarter of Quebec plants have an uptake rate under 30.32 while for Mississipi it is 13.88.
Finally, for 75% of Mississippi plants, the uptake rate is less than 28.05 whereas for Quebec it is less than 40.15.

So it appears that Quebec plants seem to have higher uptake rates than Mississippi plants.

By grouping plants by their treatment (chilled - nonchilled), we can see some characteristics: 

```{r tapply2}
tapply(CO2$uptake, CO2$Treatment, summary)
```
Plants non chilled have an uptake rate between 10.60 and 45.50  umol/m² sec whereas for chilled, it is between 7.70 and 42.40 umol/m² sec.
Moreover, for non chilled, half of plants have an uptake rate greater than 31.30 while for chilled, it is 19.70.
A quarter of non chilled plants have an uptake rate under 26.48 while for chilled it is 14.53.
Finally, for 75% of chilled plants, the uptake rate is less than 34.90 whereas for non chilled it is less than 45.50.

So it appears that non chilled plants seem to have higher uptake rates than chilled plants.


Now let's combine our two characteristics, type and treatment and take a look on two plants from Quebec, one chilled and one non chilled:

```{r tapply3}
tapply(CO2$uptake, CO2$Plant, summary)
```

Let's take the examples of Qn3 and Qc3:

- Qn3 (third plant species of Quebec nonchilled)
have an uptake rate between 16.20 and 45.50 with an average of 37.61.
Half of the Qn3 have a rate less than 42.10.
In those plants, 25% have an uptake rate less than 36.35.
Moreover, 75% have a rate less than 43.40.
In fact 50% of the population uptake rate is between 36.35 and 43.40 (mL/L).

- Qc3 (third plant species of Quebec chilled):
have an uptake rate between 15.10 and 41.40 with an average of 32.59.
Half of the Qc3 have a rate less than 38.10.
In those plants, 25% have an uptake rate less than 27.50.
Moreover, 75% have a rate less than 39.25.
In fact 50% of the population uptake rate is between 27.50 and 39.25 (mL/L).

So we see that for two plants numbered 3 from the same origin (Quebec) but one chilled and one non chilled, we obtain a significant difference in uptake rate.

We can sum up these differences using the summarize function from dplyr library.

```{r summarize}
library(dplyr)
summarize(group_by(CO2, Type), average_uptake = mean(uptake))
summarize(group_by(CO2, Treatment), average_uptake = mean(uptake))
summarize(group_by(CO2, conc), average_uptake = mean(uptake))
```

From these results, we observe that the average uptake for plants from Quebec is much higher than the one for plants from Mississippi (33.5 and 20.9 umol/m² sec respectively).
If we compare them according to the treatment, chilled plants have a smaller average uptake compared to nonchilled ones (23.8 as opposed to 30.6 umol/m² sec).
Finally, it makes sense to see that when we increase the concentration of ambient carbon dioxide, the plants tend to have higher carbon dioxide uptake rates.

## Data visualization

### Global overview

To begin with, let's just plot the data without any filtering:

```{r}
library(ggplot2)
qplot(conc, uptake, data = CO2, geom = "point",
      main = "CO2 uptake rate according to ambient CO2 concentration",
      xlab = "CO2 concentration (mL/L)", ylab = "CO2 uptake rate (umol/m² sec)")
```

The only thing we can say is that CO2 uptake rates seem to increase as ambient CO2 concentration increase.

Then, we can separate plants by their characteristics to get some insights:

```{r}
qplot(conc, uptake, facets = Treatment ~ ., data = CO2, color = Type,
      geom = c("point", "smooth"), se = FALSE,
      main = "CO2 uptake according to ambient CO2 concentration",
      xlab = "CO2 concentration (mL/L)", ylab = "CO2 uptake rate (umol/m² sec)")
```

Here we can observe several interesting things:

- As we noticed in the previous part, plants from Quebec (in red) almost always have a higher CO2 uptake rate compared to plants from Mississippi for the same ambient concentration. The smooth tendency lines clearly show that.
- Moreover, nonchilled plants (in the upper graph) seem to jump higher in terms of uptake rates that chilled ones.

We can say that this visualization confirms what we pointed out in the data exploration part.
Now, let's go further by analyzing the distribution of uptake rates according to the country of origin or treatment.

```{r}
qplot(uptake, data = CO2, color = Type, geom = "density",
      main = "Plants CO2 uptake distribution according to their origin",
      xlab = "CO2 uptake rate (umol/m² sec)")
qplot(uptake, data = CO2, color = Treatment, geom = "density",
      main = "Plants CO2 uptake distribution according to their treatment",
      xlab = "CO2 uptake rate (umol/m² sec)")
```

Again, these two plots confirm our hypothesis made above.
An important part of Mississippi plants have an uptake rate between 15 and 20 umol/m² sec whereas for Quebec plants the peak is at about 40 umol/m² sec.
For the treatment, a big part of chilled plants have an uptake rate between 15 and 20 umol/m² sec whereas for non chilled plants the peak is higher, at about 30 umol/m² sec.

### Specific overview 

#### Boxplot of uptake by origin

Now, we are going to use box plots to visualize the spreading of the uptake rates.
We can group our plants by types (Quebec and Mississippi):

```{r}
ggplot(CO2, aes(x=Type, y=uptake,  fill=Type)) +
  geom_boxplot(show.legend = FALSE) +
  ggtitle("Plants CO2 uptake rates depending on their origin") +
  xlab("Origin") + ylab("CO2 uptake rate (umol/m² sec)")
```

This figure confirms what we saw in the data exploration part. There is a huge difference in the uptake rates between plants from Quebec and plants from Mississippi. 
As we know 50% of the data is in the box and by visualizing those box plots, we clearly see that on average, plants from Quebec have higher uptake rates than Mississippi plants. 
Half of Quebec plants are over 37 umol/m² sec for the uptake rates while for Mississippi it is around 19 umol/m² sec.
In fact 25% of Quebec plants have less than 30 umol/m² sec for their uptake rate whereas more than 75% of Mississippi plants are under 29 umol/m² sec for their uptake rates.

#### Boxplots of uptake depending on treatment

Now, let's take a look at the importance of the treatment on those plants.
Let's compare the uptake rate depending on the treatment.

```{r}
ggplot(CO2, aes(x=Treatment, y=uptake,  fill=Treatment)) +
  geom_boxplot(show.legend = FALSE) +
  ggtitle("Plants CO2 uptake rates depending on treatment") +
  xlab("Treatment") + ylab("CO2 uptake rate (umol/m² sec)")
```

As we saw thanks to the apply function, non chilled plants seem to have higher uptake rates than chilled plants.
The difference between non chilled and chilled plants is less than between Quebec and Mississippi but is still important.
We clearly see that half non chilled plants are over 30 umol/m² sec for the uptake rates while for chilled it is around 19 umol/m² sec.
25% of non chilled plants have an uptake rates less than 26 umol/m² sec whereas 50% of chilled plants have less than 19 for their uptake rates.

#### Boxplots of uptake depending on treatment and type

If we combine our plots, we will obtain the following result:

```{r}
ggplot(CO2, aes(x=Type, y=uptake, fill=Treatment)) + geom_boxplot() +
  ggtitle("CO2 uptake rates of plants depending on origin with treatment specified") +
  xlab("Origin") + ylab("CO2 uptake rate (umol/m² sec)")
```

* For Quebec

In the previous part, we studied two plants from Quebec with different treatments and find out that there was a significant difference in their uptake rates. It appears clearly here as less than 75% of chilled plants are under 39 umol/m² sec while half of non chilled have an uptake rate of more than 39 umol/m² sec.

* For Mississippi

Those differences are even more relevant with Mississippi plants:
non chilled Mississippi plants have box where more than 25% are upper 22 umol/m² sec whereas for chilled plants, less than 75% of them can reach 19.

Thanks to those boxplots, we can clearly see that depending on the type of plants and the treatment, there are some chance to obtain very different uptake rates.


## Testing hypothesis

Let's work on hypothesis regarding the treatment and the type.

### Hypothesis on treatment 

Firstly, we make an assumption on the **treatment** of the plants. 
Our null hypothesis is that treatment on plant will not affect the uptake rates.
Our alternative hypothesis is that treatment on plants may affect the uptake rates.
We take a 99% confidence interval:

```{r}
t.test(CO2$uptake~CO2$Treatment, conf.level = 0.99)
```

Looking on p-value for treatment, which is 3.107e-3, we see that it is smaller than 0.01.
So, we can reject our null hypothesis and conclude that the treatment may make a difference on the uptake rates.

### Hypothesis on Type 

Secondly, we make an assumption on the **type** of the plants. 
Our null hypothesis is that type will not affect the uptake rates
Our alternative hypothesis is that type of plants may affect the uptake rates.
We take a 99% confidence interval:

```{r}
t.test(CO2$uptake~CO2$Type, conf.level = 0.99)
```

Looking on p-value for type, which is 4.451e-9, we see that it is smaller than 0.01.
So, we can reject our null hypothesis and conclude that the type may make a difference on the uptake rates.


### Combined our hypothesis

So we saw that independently, type and treatment may make a difference on the uptake rate.

We can combine our two hypothesis and make an assumption that the relationship between type and treatment will not make a difference on the uptake rates.
We combine our hypothesis and use an ANOVA test. 
T test method determines whether two groups are statistically different whereas ANOVA allows us to determines on three or more groups, that is why we have to use it.

```{r}
anova(lm(CO2$uptake ~ CO2$Type*CO2$Treatment))
```

Both t test and ANOVA look at the difference in means and the spread of the distributions. But ANOVA and t test don't determine the statistical significance in the same way. 
That is why, our p-value for respectively treatment and Type are quit different from the ones we obtain using t test.

With a confidence interval at 99%:
We obtain 2.378e-10 for type and 1.817e-4 for treatment for p-value. Those values are smaller than 0.01, so we can again reject the null hypothesis and conclude that type and treatment may make a difference on the uptake rates.

But looking at the p-value for relationship between treatment and type, we obtain 0.0642 which is bigger than 0.01. So, with an interval at 99%, we can't reject the null hypothesis and we conclude that the relationship between type and treatment did not seriously impact the uptake rates.

If we take an interval at 90%, we could reject the null hypothesis but it will deteriorate the accuracy.

#### Correlation
In the plots above, we noticed that the relationship between concentration and uptake was not perfectly linear.
So let's take a look on the correlation between them:

```{r}
cor.test(CO2$uptake, CO2$conc)
```

The correlation factor is under 50. We can improve it by using a log:

```{r}
cor.test(CO2$uptake, log(CO2$conc))
```

Using a log, we lower the impact of the concentration compare to the uptake and improve our linear correlation between those two values. It will be useful for our linear regression model.

## Fitting model

In this part, we are going to fit several models (logistic and linear ones) in order to predict the origin, the treatment and the CO2 uptake rate according to several variables.

### Model to predict the origin of a plant

The first model aims to predict the origin of the plant.
The predicted value is discrete and has two outcomes: Mississippi or Quebec.
In this case, it seems reasonable to fit a **logistic regression**.
We tried several combination of predictor variables and we obtained these accuracy values:

* uptake only: 72.6 %
* uptake + Treatment: 83.3 %
* uptake + Treatment + conc: 85.7 %
* uptake + Treatment + log(conc): 90.4 %

We remark that the more predictors we use, the better the accuracy is.
Moreover, using the log of the concentration improves the prediction because it has a better correlation than the concentration alone as we have seen in the previous part.

```{r}
type_model <- glm(Type ~ uptake + Treatment + log(conc), family = binomial, data = CO2)
summary(type_model)

type_probs <- predict(type_model, type = "response")
type_pred <- ifelse(type_probs > 0.5, "Mississippi", "Quebec")
table(Predicted=type_pred, Observed=CO2$Type)
mean(type_pred == CO2$Type)
```

Here, we chose the combination offering the best accuracy.
We can see that the model is wrong for only 8 observations out of 84.
The following graph allows us to visualize the model and its errors:

```{r}
type_pred_data <- data.frame(prob=type_model$fitted.values, Type=CO2$Type)
type_pred_data <- type_pred_data[order(type_pred_data$prob), ]
type_pred_data$rank <- 1:nrow(type_pred_data)

ggplot(data = type_pred_data, aes(x=rank, y=prob)) +
  geom_point(aes(color=Type), shape=4, stroke=2) +
  ggtitle("Probability that a plant is from Mississippi rather than Quebec") +
  xlab("Index") +
  ylab("Predicted probability")
```

### Model to predict the treatment of a plant

For this model, we use the same reasoning as for the previous one, except that we want to predict the treatment in this case.
The two outcomes are chilled or nonchilled.
Below are the accuracy obtained for several combinations:

* uptake only: 67.8 %
* uptake + Type: 70.2 %
* uptake + Type + conc: 76.1 %
* uptake + Type + log(conc): 79.7 %

```{r}
treat_model <- glm(Treatment ~ uptake + Type + log(conc), family = binomial, data = CO2)
summary(treat_model)

treat_probs <- predict(treat_model, type = "response")
treat_pred <- ifelse(treat_probs > 0.5, "chilled", "nonchilled")
table(Predicted=treat_pred, Observed=CO2$Treatment)
mean(treat_pred == CO2$Treatment)
```

We can see that the accuracy is worse than the one of the previous model (79.7 % and 90.4 %).
Indeed, the treatment is more difficult to predict than the origin because the difference of the means is smaller (see data visualization part).
The model is wrong for 17 observations out of 84.
This graph helps to visualize the model and errors made:

```{r}
treat_pred_data <- data.frame(prob=treat_model$fitted.values, Treatment=CO2$Treatment)
treat_pred_data <- treat_pred_data[order(treat_pred_data$prob), ]
treat_pred_data$rank <- 1:nrow(treat_pred_data)

ggplot(data = treat_pred_data, aes(x=rank, y=prob)) +
  geom_point(aes(color=Treatment), shape=4, stroke=2) +
  ggtitle("Probability that a plant was chilled rather than non-chilled") +
  xlab("Index") +
  ylab("Predicted probability")
```

### Model to predict the CO2 uptake rate of a plant

Finally, we fit a **linear regression** model to predict the CO2 uptake.
The outcome is thus a continuous value.
As before, we can try several combinations of predictors and compare the R² value this time:

* log(conc) only: 0.351
* log(conc) + Type: 0.697
* log(conc) + Type + Treatment: 0.799

```{r}
uptake_model <- lm(uptake ~ log(conc) + Type + Treatment, data=CO2)
summary(uptake_model)
```

We see that our fitted model is correct but not very good with a R² value of about 0.8. To get this value closer to 1, we would need more observations or other predictor variables to better predict the uptake.

To illustrate this model, we can input some test values and compare with the predictions:

```{r}
v1 = c(95, 250, 250, 250, 500, 1000)
v2 = c("Quebec", "Quebec", "Mississippi", "Mississippi", "Mississippi", "Quebec")
v3 = c("chilled", "nonchilled", "chilled", "nonchilled", "chilled", "nonchilled")
v4 = c(12.8, 37.4, 16.1, 27.5, 16.6, 43.1)
# v4 values obtained with summarize(group_by(CO2, Type, Treatment, conc), mean(uptake))

testData <- data.frame(conc=v1, Type=v2, Treatment=v3, uptake=v4)
testData["predicted_uptake"] <- predict(uptake_model, testData)
testData
```

The actual uptake values and the predicted ones are quite similar even if the prediction is far from being perfect. Our model and the dataset are not optimal so we are satisfied with these results.
